---
title: Bread 2015
author: Keith Hughitt
date: 2015/05/30
output:
  md_document:
    variant: markdown_github
---

Notes
-----

**Two challenges**:

1. Hidden hunger
   - Micronutrient (iron, etc.) defficiency
        - vitamin A, zinc, iodine, iron, folate
   - Dietary diversity
2. Obesity
   - Obesity as a form malnutrition that is an increasing problem in the
     developing world.
   - key indicators:
     - Overweight index
     - BMI (can calculate from height and weight)

- agricultural production -> hunger
- differences in nutrients for the same amount of calories
- Macro scale
    - challenge; how can we extract meaningful infomration from all of the
    variables in the harvest choice dataset?
- Micro scale
    - challenge: aggregate information

Load data
---------

```{r knitr_settings, include=FALSE}
library('knitr')
opts_knit$set(progress=FALSE, verbose=TRUE)
opts_chunk$set(fig.width=1080/96,
               fig.height=1080/96,
               dpi=96)
options(digits=4)
options(stringsAsFactors=FALSE)

```{r load_data, results='asis'}
library('ggplot2')
library('gplots')
library('readr')
library('dplyr')
library('knitr')
library('bpca')

# Malawi
mwi_nutrition = read_csv('./nutrition_indicators/Malawi/mwi_nutrition.csv')
mwi_anthropometry = read_csv('./nutrition_indicators/Malawi/mwi_anthropometry.csv')

# Example rows
#kable(t(mwi_anthropometry[1,]))
mwi_anthropometry[1,] %>% glimpse()

mwi_anthro_clean = mwi_anthropometry %>% 
    select(-wa_m, -wa_s, -ha_m, -ha_s, -wh_m, -wh_s)

#df2 = mwi_nutrition %>% select(tot_vita_A_cons, tot_zinc_cons, tot_fol_cons,
#                               tot_iron_

# Nutrition deficiencies
nutrition_def = mwi_nutrition %>% select(hhid, hh_weight, contains('def_'))

# Households found in both tables
print("Total number of households:")
length(unique(union(mwi_anthropometry$hhid, mwi_nutrition$hhid)))

common_households = intersect(mwi_anthropometry$hhid, mwi_nutrition$hhid)

# Remove incomplete rows
mwi_anthro_clean = mwi_anthro_clean[complete.cases(mwi_anthro_clean),]
summary(mwi_anthro_clean)

# Create combined table
mwi = merge(mwi_anthro_clean, nutrition_def, by='hhid')

mwi[mwi == 'No'] = FALSE
mwi[mwi == 'Yes'] = TRUE

# recode kcal_ae_hh (different from other logical variables)
mwi$def_kcal_ae_hh = ifelse(mwi$def_kcal_ae_hh == 1, TRUE, FALSE)

# save table
write.csv(mwi, 'mwi_combined_def.csv', quote=FALSE, row.names=FALSE)

# Gender colors
gender_colors = ifelse(mwi_anthro_clean$sex == 'Male', 'red', 'blue')

# Numeric part of anthropometry data
antho_clean_mat = as.matrix(mwi_anthro_clean %>% select(haz06, waz06, whz06, bmiz06))

# Biplot
plot(bpca(antho_clean_mat), var.factor=.5, obj.col=gender_colors)
title("Biplot")

# Heatmap
heatmap.2(antho_clean_mat, trace='none', dendrogram='row', RowSideColors=gender_colors)

# Ethiopia
#eth_nutrition = read_csv('./nutrition_indicators/Ethiopia/eth_nutrition.csv')
#eth_anthropometry = read_csv('./nutrition_indicators/Ethiopia/eth_anthropometry.csv')
```
