---
title: Bread 2015
author: Keith Hughitt
date: 2015/05/30
output:
  md_document:
    variant: markdown_github
---

Notes
-----

**Two challenges**:

1. Hidden hunger
   - Micronutrient (iron, etc.) defficiency
        - vitamin A, zinc, iodine, iron, folate
   - Dietary diversity
2. Obesity
   - Obesity as a form malnutrition that is an increasing problem in the
     developing world.
   - key indicators:
     - Overweight index
     - BMI (can calculate from height and weight)

- agricultural production -> hunger
- differences in nutrients for the same amount of calories
- Macro scale
    - challenge; how can we extract meaningful infomration from all of the
    variables in the harvest choice dataset?
- Micro scale
    - challenge: aggregate information

Load data
---------

```{r knitr_settings, include=FALSE}
library('knitr')
opts_knit$set(progress=FALSE, verbose=TRUE)
opts_chunk$set(fig.width=1080/96,
               fig.height=1080/96,
               dpi=96)
options(digits=4)
options(stringsAsFactors=FALSE)

```{r load_data, results='asis'}
library('ggplot2')
library('gplots')
library('readr')
library('dplyr')
library('knitr')
library('bpca')
library('reshape2')

# Malawi
mwi_nutrition = read_csv('./nutrition_indicators/Malawi/mwi_nutrition.csv')
mwi_anthropometry = read_csv('./nutrition_indicators/Malawi/mwi_anthropometry.csv')

# Example rows
#kable(t(mwi_anthropometry[1,]))
mwi_anthropometry[1,] %>% glimpse()

mwi_anthro_clean = mwi_anthropometry %>% 
    select(-wa_m, -wa_s, -ha_m, -ha_s, -wh_m, -wh_s)

#df2 = mwi_nutrition %>% select(tot_vita_A_cons, tot_zinc_cons, tot_fol_cons,
#                               tot_iron_

# Pull in some of the fields from the nutrition table 
nutrition_def = mwi_nutrition %>% 
    select(hhid, hh_female, hh_weight, contains('def_'), sh_oil_exp,
           sh_meat_exp, sh_milk_exp, foodvalue_totalyr)

# Households found in both tables
print("Total number of households:")
length(unique(union(mwi_anthropometry$hhid, mwi_nutrition$hhid)))

common_households = intersect(mwi_anthropometry$hhid, mwi_nutrition$hhid)

# Remove incomplete rows
mwi_anthro_clean = mwi_anthro_clean[complete.cases(mwi_anthro_clean),]
summary(mwi_anthro_clean)

# Create combined table
mwi = merge(mwi_anthro_clean, nutrition_def, by='hhid')

mwi[mwi == 'No'] = FALSE
mwi[mwi == 'Yes'] = TRUE

# recode kcal_ae_hh (different from other logical variables)
mwi$def_kcal_ae_hh = ifelse(mwi$def_kcal_ae_hh == 1, TRUE, FALSE)

dec_types = colnames(mwi)[grepl('def_', colnames(mwi))]
for (x in dec_types) {
    mwi[[x]] = as.logical(mwi[[x]])
}

# Gender colors
gender_colors = ifelse(mwi_anthro_clean$sex == 'Male', 'red', 'blue')

# Create columns corresponding to gender ratio in each household
hh_gender_composition = mwi %>%
    group_by(hhid) %>% 
    summarise(num_males=sum(sex == 'Male'), 
              num_females=sum(sex == 'Female')) %>%
    mutate(ratio_female = num_females / (num_females + num_males))

mwi = merge(mwi, hh_gender_composition)

# most common missing micronutrients
colSums(mwi%>% select(contains('def_')))

# Add a column: missing one or more key micronutrients: Iron, Zinc, Folate, Vitamin A
mwi = mwi %>% mutate(missing_micronutrient=(def_iron_rec_hh == TRUE     | 
                                            def_fol_rec_hh == TRUE      | 
                                            def_calcium_rec_hh == TRUE  | 
                                            def_thia_rec_hh == TRUE     | 
                                            def_vita_A_rec_hh == TRUE   | 
                                            def_vita_B6_rec_hh == TRUE  | 
                                            def_vita_B12_rec_hh == TRUE | 
                                            def_vita_C_rec_hh == TRUE   | 
                                            def_vita_E_rec_hh == TRUE   |
                                            def_zinc_rec_hh == TRUE     |
                                            def_ribof_rec_hh == TRUE
                                            ))

# Numeric part of anthropometry data
antho_clean_mat = as.matrix(mwi_anthro_clean %>% select(haz06, waz06, whz06, bmiz06))
```

Visualization
--------------

### Exploratory data analysis

#### Antropometric data

##### Biplot

```{r biplot}
plot(bpca(antho_clean_mat), var.factor=.5, obj.col=gender_colors)
title("Biplot")
```
##### Heatmap

###### Anthro data

```{r heatmap}
# Heatmap
heatmap.2(antho_clean_mat, trace='none', dendrogram='row', RowSideColors=gender_colors)
```

###### Micronutrients

```{r heatmap_extended, fig.width=1920/96, fig.height=1920/96}
gender_colors = ifelse(mwi$sex == 'Male', 'red', 'blue')

x = mwi %>% select(contains('def_'))
x[x==TRUE] = 1
x[x==FALSE] = 0
x = as.matrix(x)
heatmap.2(x, trace='none', RowSideColors=gender_colors)
```

### Nutrition

#### BMI vs. gender

```{r bmi_vs_gender}
ggplot(mwi, aes(bmiz06, fill=sex)) + geom_density(alpha=0.75)
```

#### BMI relationship to household gender composition?

```{r bmi_vs_household_gender_composition, fig.width=1920/96}
table(mwi$ratio_female)

# BMI by household gender composition (only use ratios for which we have > 5
# data points)
ggplot(mwi %>% filter(ratio_female %in% c(0, 0.5, 0.75)), 
                                aes(def_kcal_req_hh, fill=factor(ratio_female))) + 
    geom_bar(position='dodge')
```

### Female head of household

```{r female_head_of_household}
# Female head of household
ggplot(mwi, aes(def_kcal_req_hh, fill=factor(hh_female))) + 
    geom_bar(position='dodge')
```

### Micronutrient deficiency by gender of head of household

#### Individual micronutrients (and calories)

```{r deficiencies_by_head_of_household, fig.width=1920/96}
# Deficiencies
mwi_defs = mwi %>% select(hhid, hh_female, contains('def_'))
mwi_defs = melt(mwi_defs, id.vars=c('hhid', 'hh_female'))
colnames(mwi_defs) = c('hhid', 'hh_female', 'deficiency', 'is_deficient')

# Break up and plot
for (group in split(dec_types, ceiling(seq_along(dec_types) / 5))) {
    plt = ggplot(mwi_defs %>% filter(deficiency %in% group), 
           aes(is_deficient, fill=factor(hh_female))) + 
        geom_bar(position='dodge') + facet_grid(.~deficiency) +
        theme(axis.text.x=element_text(angle=-90, hjust = 0))
    print(plt)
}
```

#### Aggregate micronutrients

```{r deficiencies_by_head_of_household_agg}
table(mwi$missing_micronutrient)

ggplot(mwi, aes(missing_micronutrient, fill=hh_female)) +
    geom_bar(position='dodge') 

mwi %>% group_by(missing_micronutrient) %>%
    summarise(ratio_female=sum(hh_female == TRUE) / n())

ggplot(mwi, aes(hh_female, fill=missing_micronutrient)) +
    geom_bar(position='dodge') 

mwi %>% group_by(hh_female) %>%
    summarise(ratio_missing=sum(missing_micronutrient == TRUE) / n())

# save table
write.csv(mwi, 'output/mwi_combined.csv', quote=FALSE, row.names=FALSE)

```

### Gender / food groups

```{r gender_food_groups}
ggplot(mwi, aes(missing_micronutrient, sh_meat_exp, color=hh_female)) + 
    geom_point() +
    geom_point(position="jitter")
```

